Computing in the Field of Rationals

### 2012-06-23 rational_ext.tex
frac semantics

Provide a semantics with the non-determinism monad and define an
\emph{eval}.

Show that ``no division by zero'' is supported by the semantics.

>>>>

The operational semantics for {{PiEE*}} involve the addition of the
following two rules to the small-step operational semantics of {{Pi}}.

%subcode{opsem} include main
%! columnStyle = rclr
% <eta*, (), C> &|-->& forall v:b.{[eta*, (1/v, v), C]} & (eta*)
% <eps*, (1/v,v), C> &|-->& {[eps*, (), C]} & (eps*)

The rule {{(eta*)}} is different from other rules encountered before
due to the {{forall v:b}} quaitification and is to be understood as
follows: when the machine encounters an {{eta*}} reduction it
non-deterministically replicates into several machine states, each of
which is seeded by a unique value {{v}} that inhabits the type {{b}}
that indexes {{eta* : 1 <-> 1/b * b}}. As an example at the type
{{bool}}, we would have two possible machine states:

%subcode{opsem} include main
%! columnStyle = rclr
% <eta*, (), C> &|-->& {[eta*, (1/true, true), C]}
% and  &|-->& {[eta*, (1/false, false), C]}

The rule {{(eps*)}} is straightforward : when given a tuple
{{(1/v,v):1/b * b}} where {{eta*:1/b*b <-> 1}} the result is
{{()}}. The machine is undefined for cases where where the input is
{{(1/v',v):1/b*b}} and {{v' /= v}}.
